\documentclass{article}
\usepackage{v-test-paper}

\begin{document}
\textbf{Instructions:} For each of the following items, two statements are given—one labelled Assertion (A) and the other labelled Reason (R). Examine these two statements carefully and decide if the Assertion (A) and the Reason (R) are individually true and if the Reason (R) is the correct explanation for the Assertion (A). 

\begin{enumerate}
    \item \textbf{Assertion (A):} Insects visit flowers to gather honey. \\
    \textbf{Reason (R):} Attraction to flowers prevents the insects from damaging other parts of the plant.

    \item \textbf{Assertion (A):} In some species of asteraceae and poaceae, seeds are formed without fertilisation. \\
    \textbf{Reason (R):} Formation of fruit without fertilization is called parthenocarpy.

    \item \textbf{Assertion (A):} Endothecium layer of anther wall plays an important role in dehiscence of anther. \\
    \textbf{Reason (R):} The presence of fibrous bands and differential expansion of inner and outer tangential walls of endothecial cells cause dehiscence of anther.

    \item \textbf{Assertion (A):} Zona pellucida disappears when blastocyst reaches the uterus. \\
    \textbf{Reason (R):} Role of zona pellucida is to check the implantation of the blastocyst at an improper site.

    \item \textbf{Assertion (A):} Upto morula stage, the cells divide without any increase in size. \\
    \textbf{Reason (R):} Zona pellucida remains intact till cleavage is complete.

    \item \textbf{Assertion (A):} In a woman after hysterectomy (removal of uterus), the ovarian cycle is stopped. \\
    \textbf{Reason (R):} Stoppage of FSH secretion.

    \item \textbf{Assertion (A):} In morula stage, the cell divides without increases in size. \\
    \textbf{Reason (R):} Zona pellucida remains till cleavage.

    \item \textbf{Assertion (A):} Holoblastic cleavage, with almost equal sized blastomeres is a characteristic of placental mammals. \\
    \textbf{Reason (R):} Eggs of most mammals, including humans are of centrolecithal type.

    \item \textbf{Assertion (A):} In humans, the gamete contributed by the male determines whether the child produced will be male or female. \\
    \textbf{Reason (R):} Sex in humans is a polygenic trait depending upon a cumulative effect of some genes on X-chromosome and some on Y-chromosome.
\end{enumerate}

\begin{enumerate}
    \setcounter{enumi}{9}
    \item \textbf{Assertion}: In India, IUDs like placement of copper-T is one of most widely accepted method of contraception. \\
    \textbf{Reason}: Sterilisation procedure in the male is called vasectomy.
    
    \item \textbf{Assertion}: Sex education in schools is not necessary. \\
    \textbf{Reason}: Sex education may increase certain myths and conceptions in the students.
    
    \item \textbf{Assertion}: Now a days, there are less childless couples. \\
    \textbf{Reason}: A number of measures are now available by which even infertile couples can have child.
    
    \item \textbf{Assertion}: Over population causes a number of socio-economic problems. \\
    \textbf{Reason}: A number of bacterial and viral diseases have controlled.
    
    \item \textbf{Assertion}: Phenylketonuria is a recessive hereditary disease caused by failure of the body to oxidise an amino acid phenylalanine to tyrosine because of defective enzyme. \\
    \textbf{Reason}: It results in presence of phenylalanine in urine.
    
    \item \textbf{Assertion}: A geneticist crossed two plants. He got 50\% tall and 50\% dwarf plants in the progeny. \\
    \textbf{Reason}: One parent was heterozygous tall while the other was dwarf.
    
    \item \textbf{Assertion}: Hb$^S$ Hb$^S$, is homozygous condition of sickle cell anaemia. \\
    \textbf{Reason}: It occurs due to substitution of glutamic acid by valine at 6$^th$ position of $\beta$-chain of haemoglobin.
    
    \item \textbf{Assertion}: A middle aged woman is having small sized breasts and undersized uterus. \\
    \textbf{Reason}: Her genotype shows XO condition of allosomes.
    
    \item \textbf{Assertion}: Only a boy child could be born with the substitution of glutamic acid by valine on 6\textsuperscript{th} position of $\beta$-chain of haemoglobin. \\
    \textbf{Reason}: The gene for the above mutation occurs in Y-chromosome.
    
    \item \textbf{Assertion}: In a pedigree analysis, $\circ$ represents five unaffected offspring. \\
    \textbf{Reason}: In the pedigree analysis, the offspring are numbered with arabic numerals (1, 2, 3......) and a generation is numbered with roman numerals.
    
    \item \textbf{Assertion}: XX-XY type of sex determination mechanism is an example of male heterogamety. \\
    \textbf{Reason}: In birds, male heterogamety is observed as males produce two different types of gametes.
    
    \item \textbf{Assertion}: Number of chromosomes in one genome is equal to number of linkage groups. \\
    \textbf{Reason}: Linkage groups give important information about location of genes in the chromosomes.
    
    \item \textbf{Assertion}: Persons suffering from haemophilia fail to produce blood clotting factor VIII. \\
    \textbf{Reason}: Prothrombin producing platelets are in very low concentration.
    
    \item \textbf{Assertion}: Replication and transcription occur in the nucleus but translation takes place in the cytoplasm. \\
    \textbf{Reason}: mRNA is transferred from the nucleus into cytoplasm where ribosomes and amino acids are available for protein synthesis.
\end{enumerate}

\begin{enumerate}
    \setcounter{enumi}{23}
    \item \textbf{Assertion}: One codon may code or more than one amino acid. \\
    \textbf{Reason}: A codon is degenerate and ambiguous.

    \item \textbf{Assertion}: The first living organisms were heterotrophs. \\
    \textbf{Reason}: They were surrounded by preformed organic molecules which they used as food.

    \item \textbf{Assertion}: New life comes only from the preexisting life. \\
    \textbf{Reason}: Spontaneous generation of life under the present environmental conditions on earth is not possible.

    \item \textbf{Assertion}: Life originated by chance coming together of necessary chemicals through a series of chemical reactions (abiogenesis) \\
    \textbf{Reason}: Abiogenesis has not been experimentally proved.

    \item \textbf{Assertion}: An antibody is a protein molecule made by the lymphocytes. \\
    \textbf{Reason}: An antibody binds to a specific foreign antigen and neutralizes its odd effects.

    \item \textbf{Assertion}: Phagocyte cells digest microbes and debris. \\
    \textbf{Reason}: Natural Killer cells destroy virus-infected cells and tumour cells.

    \item \textbf{Assertion}: Lymphocytes arise from bone marrow and present in the blood, lymph and serve as natural killer cells. \\
    \textbf{Reason}: Lymphocytes migrate to thymus, where they develop into T cells and begin to mature.

    \item \textbf{Assertion}: Secondary immune response is quicker and stronger than the primary one. \\
    \textbf{Reason}: Memory cells conversant with the recombing antigens are ready to combat the invaders.

    \item \textbf{Assertion}: Sustainable agriculture aims to provide food and livelihood for present and future generations. \\
    \textbf{Reason}: Different countries and societies are observing strict environmental standards in developing advance technologies in the field of agriculture.

    \item \textbf{Assertion}: Use of fertilizers greatly enhances crop productivity. \\
    \textbf{Reason}: Irrigation is very important in increasing crop productivity.

    \item \textbf{Assertion}: Nitrogen-fixing bacteria in legume root nodules survive in oxygen-depleted cells of nodules. \\
    \textbf{Reason}: Leaghemoglobin completely removes oxygen from the nodule cells.

    \item \textbf{Assertion}: While working on staphylococci, Alexander Fleming observed that \textit{Penicillium notatum} inhibits the growth of bacteria. \\
    \textbf{Reason}: The inhibiting chemical was commercially extracted and its full potential was established by Alexander Fleming.

    \item \textbf{Assertion}: \textit{Saccharomyces cerevisiae} produces acetic acid. \\
    \textbf{Reason}: \textit{Trichoderma polysporum} produces blood cholesterol lowering agent.
\end{enumerate}

\begin{enumerate}
    \item[\textbf{37.}] \textbf{Assertion:} Curdling is required in the manufacture of cheese. \\
    \textbf{Reason:} Lactic acid bacteria are used for the purpose.
    
    \item[\textbf{38.}] \textbf{Assertion:} Endomycorrhiza of forest trees contribute to the efficient nutrient cycling in tropical forest ecosystems. \\
    \textbf{Reason:} The fungi that form mycorrhizal association with plants make nutrient ions available to them.
    
    \item[\textbf{39.}] \textbf{Assertion:} HIV is called a retro virus. \\
    \textbf{Reason:} HIV can synthesize new genomic RNA on the RNA template itself without requiring a DNA intermediate.
    
    \item[\textbf{40.}] \textbf{Assertion:} Conjugation in bacteria is not considered sexual reproduction though there is transfer of genetic material in it. \\
    \textbf{Reason:} In conjugation, meiosis, gamete formation and gametic fusion do not occur.
    
    \item[\textbf{41.}] \textbf{Assertion:} Sharbati Sonora was developed from Maxican wheat variety. \\
    \textbf{Reason:} It becomes possible by hybridization process.
    
    \item[\textbf{42.}] \textbf{Assertion:} Insulin is said to be anabolic hormone. \\
    \textbf{Reason:} Failure of insulin secretion causes diabetes.
    
    \item[\textbf{43.}] \textbf{Assertion:} Intercropping checks the population of insects. \\
    \textbf{Reason:} Plant pests can be controlled biologically by their natural parasites and pathogens.
    
    \item[\textbf{44.}] \textbf{Assertion:} Interferons are effective against viruses. \\
    \textbf{Reason:} Proteins which can be synthesized only by genetic engineering are effective against viruses.
    
    \item[\textbf{45.}] \textbf{Assertion:} Dr. Swaminathan has major contribution in breeding of wheat. \\
    \textbf{Reason:} He developed ``Sharbati Sonora'' from ``Sonora 64''.
    
    \item[\textbf{46.}] \textbf{Assertion:} Biotic potential is realised only when the environmental conditions are limiting. \\
    \textbf{Reason:} Under such conditions only, the population size can increase at the maximum rate.
    
    \item[\textbf{47.}] \textbf{Assertion:} Distribution of age group is said to influence the population growth. \\
    \textbf{Reason:} Population growth is a measure of increase in population over a period of time.
    
    \item[\textbf{48.}] \textbf{Assertion:} The relationship between sucker fish and shark is considered to be an example of parasitism. \\
    \textbf{Reason:} Sucker fish gets food and shelter from shark and also alters the growth of shark.
    
    \item[\textbf{49.}] \textbf{Assertion:} Predation and parasitism are considered to be negative interactions. \\
    \textbf{Reason:} Parasites and predators limit the population of their host species.
    
    \item[\textbf{50.}] \textbf{Assertion:} Generally the intraspecific competition is more intense than interspecific competition. \\
    \textbf{Reason:} Intraspecific competition occurs when the resources are in short supply.
\end{enumerate}

\begin{enumerate}
    \setcounter{enumi}{50}
    \item \textbf{Assertion.} Plasmodium is an endoparasite. \\
    \textbf{Reason.} Female mosquito is an example of temporary parasite.
    \item \textbf{Assertion.} Cuscuta is an example of holoparasite. \\
    \textbf{Reason.} It does not depend on other plants for nutrition requirement.
    \item \textbf{Assertion.} E. coli lives inside the human body. \\
    \textbf{Reason.} It is an example of endoparasite.
    \item \textbf{Assertion.} Cold blooded organisms show hibernation or aestivation. \\
    \textbf{Reason.} Extremes of temperature cause inactivity of organisms.
    \item \textbf{Assertion.} People living at high altitudes have higher lung capacity and more red blood corpuscles. \\
    \textbf{Reason.} Hilly persons lead a strenuous life style.
    \item \textbf{Assertion.} Cold blooded animals have no fat layer. \\
    \textbf{Reason.} Cold blooded animals use their fat for metabolic processes during hibernation.
    \item \textbf{Assertion.} Most marine animals find it difficult to live in fresh water and vice versa. \\
    \textbf{Reason.} Some animals can tolerate a narrow salinity range while others can bear a wide salinity range.
    \item \textbf{Assertion.} Frog can change colour according to its surroundings. \\
    \textbf{Reason.} Its mimicry to capture prey.
    \item \textbf{Assertion.} Interspecific competition is the only potent force in organic evolution. \\
    \textbf{Reason.} Unexceptionally two closely related species competing for the same resources cannot coexist indefinitely.
    \item \textbf{Assertion.} Presence of pneumatophores in a special adaptation of hydrophytes. \\
    \textbf{Reason.} Pneumatophores are positively geotrophic shoots that have lenticels and help in gaseous exchange.
    \item \textbf{Assertion.} shallow water rooted plants are called phytoplankton. \\
    \textbf{Reason.} They are the dominant producers in shallow waters.
    \item \textbf{Assertion.} The amount of minerals present in the soil is called standing crop. \\
    \textbf{Reason.} The growth of plants depends upon the amount of minerals available from soil.
    \item \textbf{Assertion.} Decomposition of detritus helps in biogeochemical cycling. \\
    \textbf{Reason.} Humus formed during decomposition slowly releases the minerals in the process of mineralisation.
    \item \textbf{Assertion.} Net primary productivity is less than the grass primary productivity. \\
    \textbf{Reason.} Only a part of incident energy is utilised in photosynthesis.
    \item \textbf{Assertion.} Herbivore productivity is approximately 10\% of gross productivity of producers. \\
    \textbf{Reason.} Herbivores eat less and waste a lot of food energy.
\end{enumerate}

\begin{enumerate}
    \setcounter{enumi}{65}
    \item \textbf{Assertion}: Pyramid of biomass is inverted in aquatic ecosystem. \\
    \textbf{Reason}: Longevity increases with the rise in trophic levels.
    \item \textbf{Assertion}: Primary succession occurs in a freshly harvested field. \\
    \textbf{Reason}: Succession occurs over a bare area.
    \item \textbf{Assertion}: Reed Swamp stage of hydroserre occurs in shallow water. \\
    \textbf{Reason}: The stage is characterised by the presence of amphibious plants.
    \item \textbf{Assertion}: A network of food chains existing together in an ecosystem is known as food web. \\
    \textbf{Reason}: An animal like Kite cannot be part of food web.
    \item \textbf{Assertion}: A network of food chains of an ecosystem is called food web. \\
    \textbf{Reason}: Kite cannot be part of food web.
    \item \textbf{Assertion}: In a terrestrial ecosystem, detritus food chain is a major conduit for energy flow. \\
    \textbf{Reason}: Solar energy is direct source of energy supply in detritus food chain.
    \item \textbf{Assertion}: Rice field is an ecosystem for plants and animals. \\
    \textbf{Reason}: Gut of humans/animals is an ecosystem for flora and fauna.
    \item \textbf{Assertion}: Genetic diversity within species increases with the increase in habitat variations. \\
    \textbf{Reason}: It is essential for adaptation to varied environments.
    \item \textbf{Assertion}: Hot spots are areas which have been degraded beyond repair. \\
    \textbf{Reason}: There has been adverse changes in environment.
    \item \textbf{Assertion}: Tropical rain forests are disappearing fast from developing countries such as India. \\
    \textbf{Reason}: No value is attached to these forests because these are poor in diversity.
  \end{enumerate}
\end{document}