
\item A small sphere of mass $m$ suspended by a thread is first taken aside so that the thread forms the right angle with the vertical and then released. Find:
    \begin{enumerate}
        \item the total acceleration of the sphere and the thread tension as a function of $\theta$, the angle of deflection of the thread from the vertical;
        \item the thread tension at the moment when the vertical component of the sphere's velocity is maximum;
        \item the angle $\theta$ between the thread and the vertical at the moment when the total acceleration vector of the sphere is directed horizontally.
    \end{enumerate}
\begin{solution}
    \begin{center}
        \begin{tikzpicture}
            \pic at (0, 0) {frame=3cm};
        \end{tikzpicture}
    \end{center}
    
    \begin{align*}
        \intertext{Let us depict the forces acting on the small sphere $m$, (at an arbitrary position when the thread makes an angle $\theta$ from the vertical) and write equation $\vec{F}=\vec{m}\vec{w}$ via projection on the unit vectors $\vec{t}$ and $\vec{n}$. From $F_t=mw_t$, we have}
        mg\sin\theta&=m\dfrac{dv}{dt}\\
        &=m\dfrac{vdv}{ds}=m\dfrac{vdv}{l(-d\theta)}
        \intertext{(as vertical is reference line of angular position)}
        &vdv=-gl\sin\theta d\theta
        \intertext{Integrating both the sides we get}
        \int_0^v vdv &= -gl \int_{\pi/2}^\theta \sin\theta d\theta\\
        \dfrac{v^2}{2} &= gl \cos \theta\\
        \dfrac{v^2}{l} &= 2g \cos \theta = w_n\tag{1}
        \intertext{Note: Eq. (1) can be easily obtained by the conservation of mechanical energy.}
        \intertext{From} 
        F_n &= mw_n\\
        T - mg \cos \theta &= m\dfrac{v^2}{l}
        \intertext{Using Eq. (1) we have}
        T &= 3mg \cos \theta\tag{2}
        \intertext{Again from equation $F_t = mw_t$,}
        mg \sin \theta &= mw_t \quad \text{or} \quad w_t = g \sin \theta\tag{3}
        \intertext{Hence,}
        w &= \sqrt{w_t^2 + w_n^2} = \sqrt{(g \sin \theta)^2 + (2g \cos \theta)^2} \quad \text{(using Eqs. 1 and 3)}\\
        &= g \sqrt{1 + 3 \cos^2 \theta}
        \intertext{(b) Vertical component of velocity, $v_y = v \sin \theta$}
        v_y^2 &= v^2 \sin^2\theta = 2gl \cos \theta \sin^2 \theta \quad \text{(using Eq. 1)}
        \intertext{For maximum $v_y$ or $v_y^2$,}
        \dfrac{d(\cos \theta \sin^2 \theta)}{d \theta} &= 0
        \intertext{which yields}
        \cos \theta &= \dfrac{1}{\sqrt{3}}
        \intertext{Therefore from Eq. (2),}
        T &= 3mg \dfrac{1}{\sqrt{3}} = \sqrt{3} mg
        \intertext{(c) We have $\vec{w} = \vec{w_t} + \vec{w_n}$ thus $w_y=w_{t(y)} + w_{n(y)}$}
        \intertext{But in accordance with the problem $w_y = 0$}
        \intertext{So,}
        w_{t(y)} + w_{n(y)} &= 0\\
        -g \sin \theta \sin \theta + 2g \cos \theta \cdot 0 &= 0\\
        \cos \theta &= \dfrac{1}{\sqrt{3}} \quad \text{or} \quad \theta = 54.7^\circ
    \end{align*}
\end{solution}